---
title: "Introduction to classyfireR"
author:
  - name: Thomas Wilson
    affiliation: Institute of Biological, Environmental & Rural Sciences (IBERS), Aberystwyth University, UK
    email: tpw2@aber.ac.uk
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_to_classyfireR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

[ClassyFire](http://classyfire.wishartlab.com/) is a web-based application for automated structural classification of chemical compounds. 

The `classyfireR` R package provides access to the ClassyFire RESTful API for retrieving existing compound classifications and submitted structures to the web-server for classification. 

## Installation

classyfireR can be installed from CRAN or, for the latest development version, directly from GitHub using the `remotes` package.

```{r install_cran, echo=TRUE, eval=FALSE}
install.packages('classyfireR')

remotes::install_github('aberHRML/classyfireR')
```


## Retrieving Classifications

To retrieve classifications that are already available simply provide an InChI key to the `get_classification` function. 


```{r GetClassification}
library(classyfireR)

Classification <- get_classification('BRMWTNUJHUMWMS-LURJTMIESA-N')

Classification

```

The result of each classification is stored in a single S4 (ClassyFire) object. To retrieve multiple classification, simply iterate over a vector of InChI Keys'

```{r GetClassification_Map}
InChI_Keys <-
  c('BRMWTNUJHUMWMS-LURJTMIESA-N',
    'MDHYEMXUFSJLGV-UHFFFAOYSA-N',
    'MYYIAHXIVFADCU-QMMMGPOBSA-N')


Classification_List <- purrr::map(InChI_Keys, get_classification)

```


## Submit Multiple Queries

For classification submission using SMILES, this can be performed by supplying multiple SMILES to the `submit_query` function. The results from all of the inputs, will be returned to a single S4 `Query` class. 

```{r Query}
Input <- c(MOL1 = 'CCCOCC', MOL2 = 'CNC(CC1=CN=CN1)C(=O)O')

Query <-
  submit_query(label = 'query_test',
               input = Input,
               type = 'STRUCTURE')



Query

```

## Accessor Methods

There are a series of accessor methods which will work with either object type to return results from a specific slot in the object. 

```{r Accessor}

classification(Classification)
classification(Query)


meta(Classification)
meta(Query)


chebi(Classification)
chebi(Query)

```

## Acknowledgements

If you use `classyfireR` you should cite the [ClassyFire](https://jcheminf.biomedcentral.com/articles/10.1186/s13321-016-0174-y) publication

> ___Djoumbou Feunang Y, Eisner R, Knox C, Chepelev L, Hastings J, Owen G, Fahy E, Steinbeck C, Subramanian S, Bolton E, Greiner R, and Wishart DS___. ClassyFire: Automated Chemical Classification With A Comprehensive, Computable Taxonomy. Journal of Cheminformatics, 2016, 8:61.

> __DOI:__ [10.1186/s13321-016-0174-y](https://jcheminf.biomedcentral.com/articles/10.1186/s13321-016-0174-y)

